The Elements of Theses

### 2015-03-06 thesis/InlineDedup.tex

\section{System Implementation and Experimental Evaluations}
\label{inline:impl}
We have implemented the snapshot deduplication scheme on the Aliyun's cloud platform.
Objectives of our experimental evaluation are:
1) Analyze the popularality of content data blocks and the popularity  of hot items. 
2) Assess the effectiveness  of 3-level deduplication for reducing the storage cost of snapshot 
backup. 
3) Examine the impacts of PDS size on deduplication ratio.

@@

Because there are a limited number of OS releases, we study popular blocks among
OS disks loaded with  the same OS release. 
%The selection of PDS is completed by separating OS disk and data disk.
%While block popularality among data disks may be affected in some degree
% when we scale up the number of total VMs  to be supported,
%we expect that  the block popularality among OS disks still remains to be high and  
%%can be well covered by a fixed PDS because users still adopt popular OS releases.
%we choose 7 major OSes in Aliyun's
%VM cloud platform and they are: 
%Win2008 Server 64 bits, Win2003 Server 32 bits, Win2003 Server 64 bits, RHEL, CentOS, Ubuntu Server and Debian (all Linux
%distributions are 64 bits).
%we examine 5 VM user disks from each OS, and 10 snapshots for each VM. These VMs have been actively used by their
%owners. Some of  OS disks are modified frequently and in some cases,  users even store a large amount of user data on
%their OS disks. 
%For example, some Ubuntu OS disks have only 2GB of data while some have 50GB.
%For each OS, the base image and all OS
%disk snapshots are divided into variable-sized blocks. Then 
In  Figure~\ref{fig:OSunchanged},  we list a conservative  popularality study in 7 major OS versions  supported
in Aliyun.
%We consider the earlier snapshot for each OS disk as the base image.
For every block in the base image, we classify this block as ``unchanged''
if this block has appeared in all snapshots of the same OS release even they are used by different VMs.
Figure \ref{fig:OSunchanged} shows that for each
OS,  at least 70\% of OS blocks are completely unchanged among all snapshots of the same OS. 
Some latest release  of OS tends to have a higher percentage of content change
while  old release tends to have more variations of content versions.
That can be interpreted as that users with very old  version of operating systems
have a lower tendency to update their OS versions and this causes a larger discrepancy
among OS snapshots of these users.

@@

When we deal with a large cluster of 1,000 nodes, we expect that using
1\% of data disks can cover more than what we have seen from this
1323 VM dataset, assuming that the average behavior of every subcluster with 100 machines
exhibits a similar popularality. 
In addition to this, PDS for OS disks will become even more effective when
there are more VMs sharing the same collection of OS releases. 
%Base on above data we can estimate the size of data PDS and its effect. 
%Currently we prepared 500MB memory per machine to store PDS meta, then it can represent 50GB of data. 
%If we assume each VM has 30GB of user data at runtime, and we host 25 VMs per machine, 
% maintain 10 snapshots per VM, each brings 10\% additional modified data. 
%Thus the user data in snapshot system is 1.5TB per machine. So the upper bound of 
%$PDS size/ Data size = 0.033$, which is sufficient for the 75\% saving goal.

@@

Duplicate  search approximation~\cite{extreme_binning09,sparseindex09,Xia2011}  has been proposed 
%in extreme bining and other sparse indexing 
to package similar content in one location, and duplicate lookup  only searches
for chunks within files which have a similar file-level or segment-level  content fingerprints.
That leads  to a smaller amount of memory usage for storing meta data in signature
lookup with a tradeoff of the reduced recall ratio.

>>>>

\section{System Implementation and Experimental Evaluations}
\label{inline:impl}
We have implemented the snapshot deduplication scheme on the Aliyun's cloud platform.
Objectives of our experimental evaluation are:
1) Analyze the popularity of content data blocks and the popularity  of hot items.
2) Assess the effectiveness  of 3-level deduplication for reducing the storage cost of snapshot
backup.
3) Examine the impacts of PDS size on deduplication ratio.

@@

Because there are a limited number of OS releases, we study popular blocks among
OS disks loaded with  the same OS release.
%The selection of PDS is completed by separating OS disk and data disk.
%While block popularality among data disks may be affected in some degree
% when we scale up the number of total VMs  to be supported,
%we expect that  the block popularality among OS disks still remains to be high and
%%can be well covered by a fixed PDS because users still adopt popular OS releases.
%we choose 7 major OSes in Aliyun's
%VM cloud platform and they are:
%Win2008 Server 64 bits, Win2003 Server 32 bits, Win2003 Server 64 bits, RHEL, CentOS, Ubuntu Server and Debian (all Linux
%distributions are 64 bits).
%we examine 5 VM user disks from each OS, and 10 snapshots for each VM. These VMs have been actively used by their
%owners. Some of  OS disks are modified frequently and in some cases,  users even store a large amount of user data on
%their OS disks.
%For example, some Ubuntu OS disks have only 2GB of data while some have 50GB.
%For each OS, the base image and all OS
%disk snapshots are divided into variable-sized blocks. Then
In  Figure~\ref{fig:OSunchanged},  we list a conservative  popularity study in 7 major OS versions  supported
in Aliyun.
%We consider the earlier snapshot for each OS disk as the base image.
For every block in the base image, we classify this block as ``unchanged''
if this block has appeared in all snapshots of the same OS release even they are used by different VMs.
Figure \ref{fig:OSunchanged} shows that for each
OS,  at least 70\% of OS blocks are completely unchanged among all snapshots of the same OS.
Some latest release  of OS tends to have a higher percentage of content change
while  old release tends to have more variations of content versions.
That can be interpreted as that users with very old  version of operating systems
have a lower tendency to update their OS versions and this causes a larger discrepancy
among OS snapshots of these users.

@@

When we deal with a large cluster of 1,000 nodes, we expect that using
1\% of data disks can cover more than what we have seen from this
1323 VM dataset, assuming that the average behavior of every subcluster with 100 machines
exhibits a similar popularity.
In addition to this, PDS for OS disks will become even more effective when
there are more VMs sharing the same collection of OS releases. 
%Base on above data we can estimate the size of data PDS and its effect. 
%Currently we prepared 500MB memory per machine to store PDS meta, then it can represent 50GB of data. 
%If we assume each VM has 30GB of user data at runtime, and we host 25 VMs per machine, 
% maintain 10 snapshots per VM, each brings 10\% additional modified data. 
%Thus the user data in snapshot system is 1.5TB per machine. So the upper bound of 
%$PDS size/ Data size = 0.033$, which is sufficient for the 75\% saving goal.

@@

Duplicate  search approximation~\cite{extreme_binning09,sparseindex09,Xia2011}  has been proposed 
%in extreme bining and other sparse indexing 
to package similar content in one location, and duplicate lookup  only searches
for chunks within files which have a similar file-level or segment-level  content fingerprints.
That leads  to a smaller amount of memory usage for storing meta data in signature
lookup with a trade-off of the reduced recall ratio.

### 2015-03-06 thesis/Introduction.tex

\chapter{Introduction}
\label{chap:intro}
\section{Problem Statement and Motivation}
\label{intro:prob}
The ubiquity of the cloud computing has resulted in the 
widespread availability of cluster-based services and applications accessible through 
the Internet. Examples include online storage services, big data analytics, and e-commerce websites.
In such a cluster-based cloud environment,
each physical machine runs a number of virtual machines as  instances of a guest operating system 
to contain different kind of user applications,
and their data is stored in virtual hard disks which are represented 
as virtual disk image files in the host operating system.
Frequent snapshot backup of virtual disk images is critical to increase  the service reliability
and protect data safty.
For example, the Aliyun cloud, which is  the largest cloud service provider by Alibaba in China, 
automatically conducts  the backup of virtual disk images to all active users every day.

The cost of supporting a large number of concurrent backup streams is high
because of the huge storage demand. 
If such VM snapshot data is plainly backed up without any duplicate reduction, storage waste would be 
extreme high. For example, a fresh Windows server 2008 installation costs 25 GB on the virtual disk,
and they are almost certianly duplicated with other Windows VM instances\cite{common07, pedal96}.
Using a separate  backup service with full deduplication support~\cite{venti02,bottleneck08}
can effectively identify and remove content duplicates among snapshots, 
but such a solution can be expensive. There is also a large amount of 
network traffic to transfer  data from the host machines to the backup facility
before duplicates are removed.

@@

\begin{itemize}
\item The design of a low-cost multi-stage parallel deduplication solution for automically batched deduplication.
\item An inline multi-level deduplication scheme with similarity-guided local detection and 
popularity-guided global detection.
\item The development and analysis of a VM-centric storage approach which considers fault isolation and integrates multiple duplicate detection strategies.
\item Fast and efficient garbage collection with approximate snapshot deletion and leak repair.
\end{itemize}

>>>>

\chapter{Introduction}
\label{chap:intro}
\section{Problem Statement and Motivation}
\label{intro:prob}
The ubiquity of the cloud computing has resulted in the 
widespread availability of cluster-based services and applications accessible through
the Internet. Examples include online storage services, big data analytics, and e-commerce websites.
In such a cluster-based cloud environment,
each physical machine runs a number of virtual machines as  instances of a guest operating system
to contain different kind of user applications,
and their data is stored in virtual hard disks which are represented
as virtual disk image files in the host operating system.
Frequent snapshot backup of virtual disk images is critical to increase  the service reliability
and protect data safety.
For example, the Aliyun cloud, which is  the largest cloud service provider by Alibaba in China,
automatically conducts  the backup of virtual disk images to all active users every day.

The cost of supporting a large number of concurrent backup streams is high
because of the huge storage demand.
If such VM snapshot data is plainly backed up without any duplicate reduction, storage waste would be
extreme high. For example, a fresh Windows server 2008 installation costs 25 GB on the virtual disk,
and they are almost certainly duplicated with other Windows VM instances\cite{common07, pedal96}.
Using a separate  backup service with full deduplication support~\cite{venti02,bottleneck08}
can effectively identify and remove content duplicates among snapshots,
but such a solution can be expensive. There is also a large amount of
network traffic to transfer  data from the host machines to the backup facility
before duplicates are removed.

@@

\begin{itemize}
\item The design of a low-cost multi-stage parallel deduplication solution for automatically batched deduplication.
\item An inline multi-level deduplication scheme with similarity-guided local detection and
popularity-guided global detection.
\item The development and analysis of a VM-centric storage approach which considers fault isolation and integrates multiple duplicate detection strategies.
\item Fast and efficient garbage collection with approximate snapshot deletion and leak repair.
\end{itemize}

### 2015-03-06 thesis/OfflineDedup.tex

\chapter{Synchronous and Parallel Approach for Offline Deduplication}
\label{chap:offline}
\section{Introduction}
\label{offline:intro}
This chapter introduces a low-cost architecture option and considers that
a backup service uses the existing cloud computing resource.
%Since most of backup data is not used in practice, system resource in such a service is not fully utilized. 
Performing deduplication adds significant  memory cost for comparison of content fingerprints. 
Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently. 
Cloud providers often wish that the backup service only consumes  small or modest resources 
%with a minimal impact to the existing cloud services.  Another challenge for backup with deduplication is 
with a minimal impact to the existing cloud services.  Another challenge is 
%that deletion of old snapshots compete for computing resource as well. That is because data dependence created 
that deletion of old snapshots compete for computing resource as well, because data dependence created 
by duplicate relationship among snapshots  adds processing complexity.

The traditional approach to deduplication is an inline approach which follows
a sequence of block reading, duplicate detection,  and non-duplicate  block write to the 
backup storage. In this chapter we introduce an synchronous and parallel approach
for offline batched deduplication. This solution is suitable for cloud which typically conduct
automatic backup of all the VMs during system's spare time.
Our key idea  is to  first perform parallel duplicate detection for VM content blocks 
among all machines before performing actual data backup. Each machine
accumulates detection requests and  then performs detection   partition by partition 
with minimal resource usage.
Fingerprint based partitioning allows highly parallel duplicate detection  and also simplifies 
reference counting management.  
% as the entire process can also be divided into a
%parition-wise  deletion.

While our synchronous offline solution accomplishes perfect deduplication efficiency while maintaining
low resource usage,
the trade-off of this approach is that the job turnaround time for each individual
VM backup task becomes longer. In addition,
every machine has to read dirty segments twice
and that some deduplication requests are delayed for staged parallel processing.
However, with our careful parallelism and buffer management,
this multi-stage detection scheme can provide a sufficient throughput for VM backup.

@@

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is
used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08}
uses  an in-memory Bloom filter and a prefetching cache for data blocks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
As discussed in Section~\ref{offline:intro},
there is no dedicated resource for deduplication in our targeted setting and low memory usage
 is required so that the resource impact to other cloud services is minimized.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011}
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}
to reduce memory requirement with a trade-off of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup
In comparison, this paper focuses on  full deduplication without approximation.
%We also take advantages of the fact that in a VM cloud environment,
%the virtual device driver can easily keep track if  large data
%segments have been modified using dirty bits and such information can avoid sending
%unmodified data segments for deduplication, which significantly saves cost.

>>>>

\chapter{Synchronous and Parallel Approach for Batched Deduplication}
\label{chap:offline}
\section{Introduction}
\label{offline:intro}
This chapter introduces a low-cost architecture option and considers that
a backup service uses the existing cloud computing resource.
%Since most of backup data is not used in practice, system resource in such a service is not fully utilized. 
Performing deduplication adds significant  memory cost for comparison of content fingerprints. 
Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently. 
Cloud providers often wish that the backup service only consumes  small or modest resources 
%with a minimal impact to the existing cloud services.  Another challenge for backup with deduplication is 
with a minimal impact to the existing cloud services.  Another challenge is 
%that deletion of old snapshots compete for computing resource as well. That is because data dependence created 
that deletion of old snapshots compete for computing resource as well, because data dependence created 
by duplicate relationship among snapshots  adds processing complexity.

The traditional approach to deduplication is an inline approach which follows
a sequence of block reading, duplicate detection,  and non-duplicate  block write to the 
backup storage. In this chapter we introduce an synchronous and parallel approach
for batched deduplication. This solution is suitable for cloud which typically conduct
automatic backup of all the VMs during system's spare time.
Our key idea  is to  first perform parallel duplicate detection for VM content blocks 
among all machines before performing actual data backup. Each machine
accumulates detection requests and  then performs detection   partition by partition 
with minimal resource usage.
Fingerprint based partitioning allows highly parallel duplicate detection  and also simplifies 
reference counting management.  
% as the entire process can also be divided into a
%parition-wise  deletion.

While our synchronous batched solution accomplishes perfect deduplication efficiency while maintaining
low resource usage,
the trade-off of this approach is that the job turnaround time for each individual
VM backup task becomes longer. In addition,
every machine has to read dirty segments twice
and that some deduplication requests are delayed for staged parallel processing.
However, with our careful parallelism and buffer management,
this multi-stage detection scheme can provide a sufficient throughput for VM backup.

@@

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content fingerprints to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is
used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during system's idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08}
uses  an in-memory Bloom filter and a prefetching cache for data blocks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
As discussed in Section~\ref{offline:intro},
there is no dedicated resource for deduplication in our targeted setting and low memory usage
 is required so that the resource impact to other cloud services is minimized.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011}
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}
to reduce memory requirement with a trade-off of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup
In comparison, this paper focuses on  full deduplication without approximation.
%We also take advantages of the fact that in a VM cloud environment,
%the virtual device driver can easily keep track if  large data
%segments have been modified using dirty bits and such information can avoid sending
%unmodified data segments for deduplication, which significantly saves cost.
